\thetitle

### 2021-01-12 paper/paper.tex
update

The \textit{action space} is composed of two decisions, the relative change of the lead times to the
currently set lead times $LT_{g} \in \{1,2,\ldots,\DDS\}$ for each product group $g$.
%
We restrict the lead time update for state $s_{t+1}$ according to the lead time $LT_{g}$ from state
$s_{t}$ by a maximum change of 1. Thus, if $LT_{g}$ is the current lead time of product group $g$
the agent can choose a lead time in $\{1,2,\ldots,\DDS\} \cap \{LT_{g}-1,LT_{g},LT_{g}+1\}$. % Put
% differently the algorithm can increase or decrease the lead time by 1 or leave it as it is, as long
% as it acts within the discrete action space given by the set $\{1,2,\ldots,\DDS\}$ for each product
% group.
A naive approach is to define one action for each combination of actions over all product groups.
Although applicable in our setup, as we only define two groups that yield \(9\) actions, it simply
does not scale.
%
\begin{figure*}[t!]
  \centering
  \input{figures/indep_agents}                      % Figure production system
  \caption{Illustration of independent agents as agent one increases the lead time}\label{fig:indep-agents}
\end{figure*}
%
Therefore, we invent and implement \textit{independent agents}, each operating on the same system
state, but choosing actions independently for the designated product group.
Figure~\ref{fig:indep-agents} illustrates the idea, where wlog. we use two agents and for simplicity
omit the reward and truncate the state to the lead times. Agent one chooses to increase the lead
time from \(3\) to \(4\). Thus, the future state, seen from agent one, depends on the probability of
agent two choosing to increase (\(\inc\)), decrease (\(\dec\)) or not change (\(\pers\)) the lead
time. Note that the Markov property the agent, that is
\(p(s_{t+1}, r_{t} \mid s_{t}, a_{t}) = p((4,\cdot), r_{t} \mid (3,3), \inc) \), is not touched by
this adaption. While for agent one the reward and future state only depends on the current state and
action taken, the corresponding probabilities for future state and reward change during the learning
process as the other agent starts to behave differently.
%
\MS[t]{independent agents, true?}This allows us to investigate the algorithms behaviour for action
solution spaces. In particular, we present the algorithm the above defined solution space not only
on the basis of product groups \(g\), but also for each product type \(p\).
That means, the state space, as well as the action space operates on a product type basis. Note that
without independent agents this is infeasible due to \(3^{6} = 729\) possible actions.

@@

\paragraph*{Reward.}
In each period the agent chooses an action which generates a reward while traversing to the next
period by simulating the production system. The rewards are the accumulated costs at the end of the
period consisting of the backorders, the WIP orders and the number of orders in the FGI\@. The
objective is to minimise the returned reward. Therefore, we adapt the algorithm by swapping all
maximisation functions with minimisations.

\paragraph*{Adapted Algorithm.}
Our implementation includes several optimisations, that help the algorithm in finding better
policies and stabilise the learning process. These are the above discussed independent agents that
operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous}, overestimating the
average reward \(\avgrew^{\pol}\) and ensuring to constantly reduce the average reward as searching
for policies with higher average costs is unnecessary. Algorithm~\ref{alg:full} depicts a detailed
view on the algorithm including the extensions and with the minimisation objective.

% N-Step
N-step RL collects a set of consecutive experiences, in our case retrieved from the replay memory,
and connects the state values starting from the latest observed experience, such that the newly
computed state value of the corresponding experience propagates to the experience of the previous
period. This allows a faster convergence, as values bootstrap over multiple periods in contrast to
just one step.
% Overestimating
The idea behind overestimating the average reward is to constantly seek for better policies than the
currently known. We overestimate the reward by subtracting \(10\%\) of the computed average reward
\(\avgrew^{\pol}\) and an exponentially smoothed average reward of the actually returned reward
values of greedy actions. Thus, if the computed and actually observed average reward coincide, the
overestimation equals \(0\).
% Rho Minimum
Preventing an increase of the average reward with \(\avgrew_{\max}^{\pol}\) reduces the search space
and provides a methodology for constantly improving the policy

\MS[t]{describe algorithm}

\MS[t]{rhoMinimum, (shared rho)?, init-phase, (gradient clipping)?}

Furthermore, we use multiple parallel worker agents that operate on the same
environment to collect a broader range of experiences, storing experiences according to their action
and uniformly sampling over the actions to ensure each action is constantly trained to prevent
catastrophic forgetting.

\MS[t]{explanation missing}

% The algorithm depicted above is designed as a tabular version. We have adapted it to use neural
% networks for the approximation of the different kind of state-action pair value functions. This is
% required due to the exponential growth of the state space. The implementation collects the observed
% information for each state-action function in a replay memory and randomly selects \(128\) data
% points out of a collection of \(30k\) memories in each period and trains these on the neural
% network. This technique is called experience replay memory
% \citep{lin1993reinforcement,mnih2015human} and ought to overcome we instabilities. Nonetheless, this
% approximation results in the fact that the algorithm can be unstable or may even diverge
% \citep{tsitsiklis1997analysis}. Additionally we use a target and a worker network, whereas the
% target network is overwritten every \(10k\) steps by the worker network. This has been proposed by
% \cite{mnih2015human} and should further stabilize the algorithm.

\MS[t]{Extensions: overestimate rho, rho minimum, Multiple Agents (shared state,  shared rho?),
  init-phase (0.5 for expSmthRewRate, always adapt rho, same rho for all agents): exp smooth reward
  as rho, replay memory actions, n-step (describe replay memory), gradient clipping }
\MS[t]{Side info: Replay memory also stores (disallowed) filtered action indices for optimisation purposes and on
episodic tasks needs to store a bool indicating an episode end. Uses Int8 for state features values.  }

\begin{algorithm}[t!]
  \begin{algorithmic}[1]
    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly,
    set an exploration rate \(0 \leqslant \plearn \leqslant \pexp \leqslant 1\), exponential
    smoothing learning rates \(0 < \alpha_{\max}, \alpha, \gamma < 1\), a sufficiently large default
    value for \(\avgrew_{\max}^{\pol}\) and discount factors \(0 < \gamma_{0} < \gamma_{1} \leqslant
    1\)
    \While{the stopping criterion is not fulfilled}
    \State{\(\isRandAct{t} \gets \mathsf{False}\)}
    \ForEach{independent agent \(\agent_{1},\agent_{2},\ldots,\agent_{n_{\mathsf{ag}}}\)}
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
      With probability \(\frac{\pexp}{n_{\mathsf{ag}}}\) choose a random action and
      probability \(1-\frac{\pexp}{n_{\mathsf{ag}}}\) one that fulfills
      \(\min_{a}\lexeq(\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{}),\X^\pol_ {\gamma_{0}}(s_{t},a;
      \netWorker{}))\). Set \(\isRandAct{t} \gets \mathsf{True}\) if the action was chosen randomly
    \end{minipage}
    \EndFor
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
    Carry out action \(a_{t}\), observe reward \(r_{t}\) and resulting state \(s_{t+1}\).
    Store the experience \((s_{t}, a_{t}, \isRandAct{t}, r_{t}, s_{t+1})\) in the experience replay
    memory \(M\).
    \end{minipage}
    \If{\(t \mod{\nstep} \equiv 0\)}
    \State{Reset gradient: \(d\netWorker{} \gets 0\)}
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
      Sample random mini-batch of \(\nstep\) consecutive experiences from \(M\), each with
      \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i})\) where \(i \in {1,2,\ldots,\nstep}\).
    \end{minipage}
    \ForEach{mini-batch}
    \ForEach{experience, starting with the latest seen experience \(\nstep\)}
    \If{experience \(\nstep\) or (\(\isRandAct{i}\) and  \(\pexp \leqslant \plearn\))}
      \State{\(\target_{\gamma0} \gets \min_{a} \X^\pol_{\gamma_0}(s_{t+1},a; \netTarget{})\)}
      \State{\(\target_{\gamma1} \gets \min_{a} \X^\pol_{\gamma_1}(s_{t+1},a; \netTarget{})\)}
    \EndIf
    \If{a non-random action was chosen or \(\pexp > \plearn\)}
    \State{\(\avgrew^{\pol} \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
      \min_{a}\X^\pol_{\gamma_1}(s_{t+1},a; \netWorker{}) - \X^\pol_{\gamma_1}(s_{t},a_{t};
      \netWorker{})]\)}
    \State{\(\avgrew^{\pol} \gets \min(\avgrew^{\pol}, \avgrew_{\max}^{\pol})\)}
    \State{\(\avgrew_{\max}^{\pol} \gets (1- \alpha_{\max}) \avgrew_{\max}^{\pol} +
    \alpha_{\max} 1.025 \avgrew^{\pol}\)}
    \EndIf
    \State{Compute overestimate of \(\avgrew^{\pol}\) and store result in \(\avgrew_{\Overestim}^{\pol}\)}
    \State{\(\target_{\gamma_0}  \gets r_{t} + \gamma_{0} \target_{\gamma_{0}} - \avgrew_{\Overestim}^{\pol}  \)}
    \State{\(\target_{\gamma_{1}}  \gets r_{t} + \gamma_{1} \target_{\gamma_{1}} - \avgrew_{\Overestim}^{\pol}\)}
    \State{\(d\netWorker{} \gets d\netWorker{} + \partial{(\target_{\gamma_{0}} - \X^\pol_{\gamma_{0}}(s_{t+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
    \State{\(d\netWorker{} \gets d\netWorker{} +
      \partial{(\target_{\gamma_{1}} - \X^\pol_{\gamma_{1}}(s_{t+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
    \EndFor{}
    \EndFor{}
    \State{Perform update of \(\netWorker{}\) using \(d\netWorker{}\)}
    \EndIf{}
    \State{Every \(C\) steps exponential smoothly set target network:
    \(\netTarget \gets (1 - \gamma) \netTarget{} + \gamma \netWorker{}\)}
    \State{Set \(s \gets s'\), \(t \gets t+1\) and decay parameters}
    \EndWhile{}
  \end{algorithmic}
  \caption{\label{alg:full}Near-Blackwell-optimal deep RL for unichain MDPs}
\end{algorithm}

>>>>

The \textit{action space} is composed of two decisions, the relative change of the lead times to the
currently set lead times $LT_{g} \in \{1,2,\ldots,\DDS\}$ for each product group $g$.
%
We restrict the lead time update for state $s_{t+1}$ according to the lead time $LT_{g}$ from state
$s_{t}$ by a maximum change of 1. Thus, if $LT_{g}$ is the current lead time of product group $g$
the agent can choose a lead time in $\{1,2,\ldots,\DDS\} \cap \{LT_{g}-1,LT_{g},LT_{g}+1\}$. % Put
% differently the algorithm can increase or decrease the lead time by 1 or leave it as it is, as long
% as it acts within the discrete action space given by the set $\{1,2,\ldots,\DDS\}$ for each product
% group.
A naive approach is to define one action for each combination of actions over all product groups.
Although applicable in our setup, as we only define two groups that yield \(9\) actions, it simply
does not scale.
%
\begin{figure*}[t!]
  \centering
  \input{figures/indep_agents}                      % Figure production system
  \caption{Illustration of independent agents as agent one increases the lead time}\label{fig:indep-agents}
\end{figure*}
%
Therefore, we invent and implement \textit{independent agents}, each operating on the same system
state, but choosing actions independently for the designated product group.
Figure~\ref{fig:indep-agents} illustrates the idea, where wlog. we use two agents and for simplicity
omit the reward and truncate the state to the lead times. Agent one chooses to increase the lead
time from \(3\) to \(4\). Thus, the future state, seen from agent one, depends on the probability of
agent two choosing to increase (\(\inc\)), decrease (\(\dec\)) or not change (\(\pers\)) the lead
time. Note that the Markov property the agent, that is
\(p(s_{t+1}, r_{t} \mid s_{t}, a_{t}) = p((4,\cdot), r_{t} \mid (3,3), \inc) \), is not touched by
this adaption. While for agent one the reward and future state only depends on the current state and
action taken, the corresponding probabilities for future state and reward change during the learning
process as the other agent starts to behave differently.
%
\MS[t]{independent agents, true?}This allows us to investigate the algorithms behaviour for even
larger solution spaces. In particular, this adaption enables us to present the algorithm the above
defined solution space not only on the basis of product groups \(g\), but also for each product type
\(p\). That means, the state space, as well as the action space operates on a product type basis.
Note that without independent agents this is infeasible due to \(3^{6} = 729\) possible actions.
Removing the dependency between the lead times of different product types reduces the possibilities,
and thus the number of ANN output layer nodes, to \(18\). Therefore, this is a vital adaption that
ensures scalability.

@@

\paragraph*{Reward.}
In each period the agents choose actions which generates a reward while traversing to the next
period by simulating the production system. The rewards are the accumulated costs at the end of the
period consisting of the backorders, the WIP orders and the number of orders in the FGI\@. The
objective is to minimise the returned reward. Therefore, we adapt the algorithm by swapping all
maximisation functions with minimisations.

\paragraph*{Adapted Algorithm.}
Our implementation includes several optimisations, that help the algorithm in better exploring the
solution space and stabilise the learning process. These include the above discussed independent
agents that operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous},
overestimating the average reward \(\avgrew^{\pol}\) and ensuring to constantly reduce the average
reward as searching for policies with higher average costs is unnecessary. Algorithm~\ref{alg:full}
depicts a detailed view on the algorithm including these extensions and with the minimisation
objective. The idea is the same as in Algorithm~\ref{alg:near}, except that the action selection
procedure is a loop over all independent agents and the implementation of the n-step logic.

% N-Step
N-step RL collects a set of consecutive experiences, in our case retrieved from the replay memory,
and connects the state values starting from the latest observed experience, such that the newly
computed state value of the corresponding experience propagates to the experience of the previous
period. This allows a faster convergence, as values bootstrap over multiple periods in contrast to
just one step.
% Overestimating
The idea behind overestimating the average reward is to constantly seek for better policies than the
currently known. We overestimate the reward by subtracting \(10\%\) of the computed average reward
\(\avgrew^{\pol}\) and an exponentially smoothed average reward of the actually returned reward
values of greedy actions. Thus, if the computed and actually observed average reward coincide, the
overestimation equals \(0\).
% Rho Minimum
Preventing an increase of the average reward with \(\avgrew_{\max}^{\pol}\) reduces the search space
and provides a methodology for constantly improving policy.

\MS[t]{(shared rho)?, (gradient clipping)?}

Furthermore, we sometimes use multiple parallel workers (parallel RL) that operate on the same
environment to collect a broader range of experiences. The experiences are stored according to the
first taken action in the n-step series. We uniformly sample over the actions to ensure each action
is constantly trained. This reduces catastrophic forgetting.

The algorithm initiation tactics is to first fill the replay memory buffer by skipping any learning
(steps \(10-25\)) until the period, that corresponds to the replay memory size, is reached. Then
for \(500\) steps the average reward is set to the last computed exponentially smoothed value to
prevent divergence due to the untrained network weight initialisation values.

% \MS[t]{Extensions: overestimate rho, rho minimum, Multiple Agents (shared state,  shared rho?),
%   init-phase (0.5 for expSmthRewRate, always adapt rho, same rho for all agents): exp smooth reward
%   as rho, replay memory actions, n-step (describe replay memory), gradient clipping }
% \MS[t]{Side info: Replay memory also stores (disallowed) filtered action indices for optimisation purposes and on
% episodic tasks needs to store a bool indicating an episode end. Uses Int8 for state features values.  }

\begin{algorithm}[t!]
  \begin{algorithmic}[1]
    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly,
    set an exploration rate \(0 \leqslant \plearn \leqslant \pexp \leqslant 1\), exponential
    smoothing learning rates \(0 < \alpha_{\max}, \alpha, \gamma < 1\), a sufficiently large default
    value for \(\avgrew_{\max}^{\pol}\) and discount factors \(0 < \gamma_{0} < \gamma_{1} \leqslant
    1\)
    \While{the stopping criterion is not fulfilled}
    \State{\(\isRandAct{t} \gets \mathsf{False}\)}
    \ForEach{independent agent \(\agent \in \agent_{1},\agent_{2},\ldots,\agent_{n_{\mathsf{ag}}}\)}
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
      With probability \(\frac{\pexp}{n_{\mathsf{ag}}}\) choose a random action and
      probability \(1-\frac{\pexp}{n_{\mathsf{ag}}}\) one that fulfills
      \(\min_{a \in \mathcal{A}(\agent)}\lexeq(\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{}),\X^\pol_ {\gamma_{0}}(s_{t},a;
      \netWorker{}))\). Set \(\isRandAct{t} \gets \mathsf{True}\) if the action was chosen randomly
    \end{minipage}
    \EndFor
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
    Carry out combined action \(a_{t}\), observe reward \(r_{t}\) and resulting state \(s_{t+1}\).
    Store the experience \((s_{t}, a_{t}, \isRandAct{t}, r_{t}, s_{t+1})\) in the experience replay
    memory \(M\).
    \end{minipage}
    \If{\(t \mod{\nstep} \equiv 0\)}
    \State{Reset gradient: \(d\netWorker{} \gets 0\)}
    \State{}%
    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
      Sample random mini-batch of \(\nstep\) consecutive experiences from \(M\), each with
      \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i})\) where \(i \in {1,2,\ldots,\nstep}\).
    \end{minipage}
    \ForEach{mini-batch}
    \ForEach{experience, starting with the latest seen experience \(\nstep\)}
    \If{experience \(\nstep\) or (\(\isRandAct{i}\) and  \(\pexp \leqslant \plearn\))}
      \State{\(\target_{\gamma0} \gets \min_{a} \X^\pol_{\gamma_0}(s_{t+1},a; \netTarget{})\)}
      \State{\(\target_{\gamma1} \gets \min_{a} \X^\pol_{\gamma_1}(s_{t+1},a; \netTarget{})\)}
    \EndIf
    \If{a non-random action was chosen or \(\pexp > \plearn\)}
    \State{\(\avgrew^{\pol} \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
      \min_{a}\X^\pol_{\gamma_1}(s_{t+1},a; \netWorker{}) - \X^\pol_{\gamma_1}(s_{t},a_{t};
      \netWorker{})]\)}
    \State{\(\avgrew^{\pol} \gets \min(\avgrew^{\pol}, \avgrew_{\max}^{\pol})\)}
    \State{\(\avgrew_{\max}^{\pol} \gets (1- \alpha_{\max}) \avgrew_{\max}^{\pol} +
    \alpha_{\max} 1.025 \avgrew^{\pol}\)}
    \EndIf
    \State{Compute overestimate of \(\avgrew^{\pol}\) and store result in \(\avgrew_{\Overestim}^{\pol}\)}
    \State{\(\target_{\gamma_0}  \gets r_{t} + \gamma_{0} \target_{\gamma_{0}} - \avgrew_{\Overestim}^{\pol}  \)}
    \State{\(\target_{\gamma_{1}}  \gets r_{t} + \gamma_{1} \target_{\gamma_{1}} - \avgrew_{\Overestim}^{\pol}\)}
    \State{\(d\netWorker{} \gets d\netWorker{} + \partial{(\target_{\gamma_{0}} - \X^\pol_{\gamma_{0}}(s_{t+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
    \State{\(d\netWorker{} \gets d\netWorker{} +
      \partial{(\target_{\gamma_{1}} - \X^\pol_{\gamma_{1}}(s_{t+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
    \EndFor{}
    \EndFor{}
    \State{Perform update of \(\netWorker{}\) using \(d\netWorker{}\)}
    \EndIf{}
    \State{Every \(C\) steps exponential smoothly set target network:
    \(\netTarget \gets (1 - \gamma) \netTarget{} + \gamma \netWorker{}\)}
    \State{Set \(s \gets s'\), \(t \gets t+1\) and decay parameters}
    \EndWhile{}
  \end{algorithmic}
  \caption{\label{alg:full}Near-Blackwell-optimal deep RL for unichain MDPs}
\end{algorithm}
